---
editor_options: 
  markdown: 
    wrap: 72
---

**The task this week is to:**

-   Read in [global gender inequality
    data](https://hdr.undp.org/data-center/documentation-and-downloads)

-   Join the global gender inequality index to [spatial data of the
    World](https://hub.arcgis.com/datasets/2b93b06dc0dc4e809d3c8db5cb96ba69_0/explore?location=-2.688200%2C0.000000%2C1.41),
    creating a new column of difference in inequality between 2010 and
    2019

```{=html}
<!-- -->
```
-   Share it with the World on GitHub

-   Add you repository URL to the circulated spreadsheet

[**Tip:**]{.underline}

-   The [countrycode](https://github.com/vincentarelbundock/countrycode)
    R package will be helpful!

-   The gender inequality has changed in the last year, you will find
    what you need in the "All composite indices and components time
    series (1990-2021)" dataset, the metadata file beneath it will
    explain what the columns are.

### Install packages and load them in

```{r}
library(tidyverse)
library(janitor)
library(terra)
library(here)
install.packages('remotes')
library(remotes)
install_github('vincentarelbundock/countrycode')
library(countrycode)
library(tmap)
```

### Read in inequality data and spatial data of the world

```{r}

world <-
  st_read(here("wk4_data", "World_Countries_Generalized.geojson"))

View(world)

??countrycode
```

```{r}
hdrdata <-
  read_csv(here("wk4_data","HDR21-22_Composite_indices_complete_time_series.csv"))

gii <- hdrdata %>%
  select(iso3,
         country,
         hdicode,
         region,
         gii_2010,
         gii_2019) %>%
  mutate(iso = countrycode(
    iso3,
    origin = 'iso3c',
    destination = 'iso2c',
    nomatch = NA
  )) %>%
  mutate(gii_diff = gii_2019 - gii_2010)
  
```

```{r}
gii
```

### Join them and add a column of difference 2010 and 2019

```{r}
gii1019_map <- world %>% 
  left_join(.,
            gii,
            by = c('ISO'='iso')
            )
```

```{r}
tmap_mode("plot")
map <- gii1019_map %>% 
  tm_shape(.) + 
  tm_fill('gii_diff') +
  tm_layout(title = 'Gender Inequality Index difference 2010-2019')
map
```

### Share on GitHub
